Performance Evaluation of Big Data Processing Strategies for Neuroimaging

### 2018-12-13 paper.tex
Started discussion

\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{colortbl} % for \rowcolor
\usepackage{caption}
\usepackage{subcaption}
\usepackage[bookmarks=false]{hyperref}
\usepackage{algorithm}
\usepackage{algorithmicx} % Doc is at http://tug.ctan.org/macros/latex/contrib/algorithmicx/algorithmicx.pdf
\usepackage{algpseudocode}
\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
\usepackage[bookmarks=false]{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,filecolor=black,urlcolor=blue}

@@

\begin{abstract}

\end{abstract}

@@

@@

Processing large images is only considered to be part of the Big Data problem 
in neuroscience. The other problem is processing large MRI datasets, that is,
datasets consisting of many small brain images belonging to various different 
subjects. Although we have not explored explicitly the processing of large MRI
datasets, the 75GB BigBrain is within the size ballpark~\cite{van2014human} of 
MRI datasets commonly processed in today's studies.

@@

\begin{table*}
\centering
\begin{tabular}{c|ccc|cccc|cc}
  \rowcolor{headcolor}
  \multicolumn{10}{c}{Experiment 1: Number of Iterations}\\
  \hline
  \rowcolor{headcolor}
 n &
 D (GB) & C (s) & D/C (MB/s) &
 $\gamma$ & $\delta/\gamma$ (MB/s) & $\Gamma$ & $\Delta/\Gamma$ (MB/s)&
 \parbox{2cm}{Page-cache on local disk? (Eq.~\ref{eq:page-cache-inequality})} & \parbox{2cm}{Page cache on Lustre? (Eq.~\ref{eq:page-cache-sharedfs})}\\
 \hline
1   &75    & 430    & 178.6 & 9 & 21.5 & 125 & 4.0 & \cellcolor{red!20} No & \cellcolor{red!20} No\\
10  & 750   & 4,300 & 178.6 & 9 & 21.5 & 125 & 4.0 & \cellcolor{red!20} No & \cellcolor{red!20} No\\
100 & 7,500 & 43,000& 178.6 & 9 & 21.5 & 125 & 4.0 & \cellcolor{red!20} No & \cellcolor{red!20} No\\
\hline
  \multicolumn{10}{c}{}\\

\rowcolor{headcolor}
  \multicolumn{10}{c}{Experiment 2: Task Duration}\\
  \hline
  \rowcolor{headcolor}
 x (s) &
 D (GB) & C (s) & D/C (MB/s) &
 $\gamma$ & $\delta/\gamma$ (MB/s) & $\Gamma$ & $\Delta/\Gamma$ (MB/s)&
 \parbox{2cm}{Page-cache on local disk? (Eq.~\ref{eq:page-cache-inequality})} & \parbox{2cm}{Page cache on Lustre? (Eq.~\ref{eq:page-cache-sharedfs})}\\
 \hline
 2.4  & 750 & 3,000   & 256   & 9 & 21.5 & 125 & 4.0 &  \cellcolor{red!20} No & \cellcolor{red!20} No   \\
 3.44 & 750 & 4,300   & 178.6 & 9 & 21.5 & 125 & 4.0 &  \cellcolor{red!20} No & \cellcolor{red!20} No   \\
 7.68 & 750 & 9,600   & 80    & 9 & 21.5 & 125 & 4.0 &  \cellcolor{red!20} No & \cellcolor{red!20} No   \\
 320  & 750 & 400,000 & 1.9   & 9 & 21.5 & 125 & 4.0 &  \cellcolor{green!20} Yes & \cellcolor{green!20} Yes  \\
 \hline
  \multicolumn{10}{c}{}\\

 \rowcolor{headcolor}
  \multicolumn{10}{c}{Experiment 3: Number of Chunks}\\
  \hline
  \rowcolor{headcolor}
 chunks &
 D (GB) & C (s) & D/C (MB/s) &
 $\gamma$ & $\delta/\gamma$ (MB/s) & $\Gamma$ & $\Delta/\Gamma$ (MB/s)&
 \parbox{2cm}{Page-cache on local disk? (Eq.~\ref{eq:page-cache-inequality})} & \parbox{2cm}{Page cache on Lustre? (Eq.~\ref{eq:page-cache-sharedfs})}\\
 \hline
 30  & 750 & 4,400   & 174.6 & 2  & 96.8 & 30  & 16.8 &  \cellcolor{red!20} No & \cellcolor{red!20} No   \\
 125 & 750 & 4,400   & 174.6 & 9  & 21.5 & 125 & 4.0 &  \cellcolor{red!20} No & \cellcolor{red!20} No   \\
 750 & 750 & 4,400   & 174.6 & 25 & 7.7  & 375 & 1.3 &  \cellcolor{red!20} No & \cellcolor{red!20} No  \\

\hline
  \multicolumn{10}{c}{}\\ 
 \rowcolor{headcolor}
  \multicolumn{10}{c}{Experiment 4: Image Size}\\
  \hline
  \rowcolor{headcolor}
 image  &
 D (GB) & C (s) & D/C (MB/s) &
 $\gamma$ & $\delta/\gamma$ (MB/s) & $\Gamma$ & $\Delta/\Gamma$ (MB/s)&
 \parbox{2cm}{Page-cache on local disk? (Eq.~\ref{eq:page-cache-inequality})} & \parbox{2cm}{Page cache on Lustre? (Eq.~\ref{eq:page-cache-sharedfs})}\\
 \hline
 BigBrain      & 750   & 2,200   & 349.1     & 9  & 21.5  & 125 & 4.0 &  \cellcolor{red!20} No & \cellcolor{red!20} No   \\
 Half BigBrain & 375   & 2,200   & 174.6     & 9  & 21.5  & 125 & 4.0 &  \cellcolor{red!20} No & \cellcolor{red!20} No   \\
 MRI           & 0.127 & 2,200   & 0.06 & 9& 21.5  & 125 & 4.0 &  \cellcolor{green!20} Yes & \cellcolor{green!20} Yes  
\end{tabular}
\caption{Experiment conditions}
\label{table:experiments}
\end{table*}

@@

\begin{figure*}
\begin{subfigure}{\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{results/figures/iterations.pdf}
    \caption{Experiment 1: complete BigBrain, 125 chunks, 3.44-second tasks.}\label{fig:iterations}
\end{subfigure}
\begin{subfigure}{\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{results/figures/cputime.pdf}
    \caption{Experiment 2: complete BigBrain, 125 chunks, 10 iterations.}\label{fig:cputime}
\end{subfigure}
\begin{subfigure}{\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{results/figures/numchunks.pdf}
    \caption{Experiment 3: complete BigBrain, 10 iterations, C= 4,400s.}\label{fig:numchunks}
\end{subfigure}
\begin{subfigure}{\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{results/figures/datasize.pdf}
    \caption{Experiment 4: 125 chunks, 10 iterations, 1.76-second tasks.}\label{fig:datasize}
\end{subfigure}
\caption{Experiment results: Makespans of Spark and Nipype writing to memory, tmpfs, local  disk and Lustre.}
\end{figure*}
\subsection{Experiment 1: Number of Iterations}

@@

\subsection{Page Cache Model Evaluations}
\begin{figure*}
\begin{subfigure}{0.5\linewidth}
\centering
    \includegraphics[width=\textwidth]{results/figures/local-incrementation.pdf}
\caption{Local Disk}
\label{fig:sub2}
    \end{subfigure}%
\begin{subfigure}{0.5\linewidth}
\centering
    \includegraphics[width=\textwidth]{results/figures/lustre-incrementation.pdf}
\caption{Lustre}
\label{fig:sub3}
\end{subfigure}
    \caption{Model evaluation. Grey 
             regions denote areas that violate model predictions.}
\label{fig:modeleval}
\end{figure*}

In order to evaluate the page cache model, we compared the observed 
speedup ratio provided by in-memory computing to the (D/C) / 
($\delta$/$\gamma$) and (D/C) / 
($\Delta$/$\Gamma$) ratios (Fig.~\ref{fig:modeleval}). Experiments for 
which there was no in-memory equivalent (i.e. BigBrain split into 30 
chunks) 
were not considered.

@@

\section{Discussion} % 2 pages
\label{sec:discussion}
\subsection{Effect of In-Memory Computing}
% Shared fs vs local disk vs in-memory: what do we gain?

\subsection{Effect of Data Locality}
% It should be important, in particular when there is contention and when
% tasks are small.

>>>>

\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{colortbl} % for \rowcolor
\usepackage{caption}
\usepackage{subcaption}
\usepackage[bookmarks=false]{hyperref}
\usepackage{algorithm}
\usepackage{algorithmicx} % Doc is at http://tug.ctan.org/macros/latex/contrib/algorithmicx/algorithmicx.pdf
\usepackage{algpseudocode}
\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,filecolor=black,urlcolor=blue}

@@

\begin{abstract}
We benchmark Big Data processing strategies, namely in-memory 
computing, data locality, and lazy evaluation, on typical neuroimaging 
processing use cases represented by the BigBrain dataset. We use Dell 
EMC's Top-500 cluster, as well as Apache Spark and Nipype as 
representative processing engines. We model Big Data thresholds by 
comparing the data/compute ratio of the application to the disk and 
shared file system bandwidths reported to the number of concurrent 
processes in a node and across the cluster. This model acknowledges the 
fact that page caching provided by the Linux kernel is critical to the 
performance of Big Data applications. Results show that in-memory 
computing alone speeds-up executions by a factor of up to 1.6, while 
combined with data locality this factor reaches 5.3. Such important 
speed-up values are likely to be observed on typical image processing 
operations performed on images of size larger than 75GB. A ballpark 
speculation from our model showed that in-memory computing alone would 
not speed-up current 
functional MRI analyses, but that it would when coupled with data 
locality, for a number of concurrently  processed subjects in the order 
of 200. In addition, we observe that emulating in-memory computing 
using in-memory file systems (tmpfs) does not reach the performance of 
an in-memory engine, presumably due to swapping to disk and the lack of data cleanup.
We conclude that Big Data processing strategies are 
worth developing for neuroimaging 
applications. 
\end{abstract}

@@

\todo{Remove if not included}

@@

Processing large images is only considered to be part of the Big Data problem 
in neuroscience. The other problem is processing large MRI datasets, 
that is, datasets consisting of many small brain images belonging to 
various different subjects. This situation is commonly observed in 
functional MRI (fMRI), where it is becoming increasingly common to 
process data from hundreds of subjects. Although we have not explored 
explicitly the processing of large MRI datasets, the 75GB BigBrain is 
within the size ballpark~\cite{van2014human} of 
MRI datasets commonly processed in today's studies.

@@

\begin{table*}
\centering
\begin{tabular}{c|ccc|cccc|cc}
  \rowcolor{headcolor}
  \multicolumn{10}{c}{Experiment 1: Number of Iterations}\\
  \hline
  \rowcolor{headcolor}
 n &
 D (GB) & C (s) & D/C (MB/s) &
 $\gamma$ & $\delta/\gamma$ (MB/s) & $\Gamma$ & $\Delta/\Gamma$ (MB/s)&
 (D/C)/($\delta/\gamma$) & (D/C)/($\Delta/\Gamma$)\\
 \hline
1   &75    & 430    & 178.6 & 9 & 21.5 & 125 & 4.0 & \cellcolor{red!20} 8.3 & \cellcolor{red!20} 44.7\\
10  & 750   & 4,300 & 178.6 & 9 & 21.5 & 125 & 4.0 & \cellcolor{red!20} 8.3 & \cellcolor{red!20} 44.7\\
100 & 7,500 & 43,000& 178.6 & 9 & 21.5 & 125 & 4.0 & \cellcolor{red!20} 8.3 & \cellcolor{red!20} 44.7\\
\hline
  \multicolumn{10}{c}{}\\

\rowcolor{headcolor}
  \multicolumn{10}{c}{Experiment 2: Task Duration}\\
  \hline
  \rowcolor{headcolor}
 x (s) &
 D (GB) & C (s) & D/C (MB/s) &
 $\gamma$ & $\delta/\gamma$ (MB/s) & $\Gamma$ & $\Delta/\Gamma$ (MB/s)&
 (D/C)/($\delta/\gamma$) & (D/C)/($\Delta/\Gamma$)\\
 \hline
 2.4  & 750 & 3,000   & 256   & 9 & 21.5 & 125 & 4.0 &  \cellcolor{red!20} 11.9 & \cellcolor{red!20} 64   \\
 3.44 & 750 & 4,300   & 178.6 & 9 & 21.5 & 125 & 4.0 &  \cellcolor{red!20} 8.3 & \cellcolor{red!20} 44.7   \\
 7.68 & 750 & 9,600   & 80    & 9 & 21.5 & 125 & 4.0 &  \cellcolor{red!20} 3.7 & \cellcolor{red!20} 20   \\
 320  & 750 & 400,000 & 1.9   & 9 & 21.5 & 125 & 4.0 &  \cellcolor{green!20} 0.09 & \cellcolor{green!20} 0.48  \\
 \hline
  \multicolumn{10}{c}{}\\

 \rowcolor{headcolor}
  \multicolumn{10}{c}{Experiment 3: Number of Chunks}\\
  \hline
  \rowcolor{headcolor}
 chunks &
 D (GB) & C (s) & D/C (MB/s) &
 $\gamma$ & $\delta/\gamma$ (MB/s) & $\Gamma$ & $\Delta/\Gamma$ (MB/s)&
 (D/C)/($\delta/\gamma$) & (D/C)/($\Delta/\Gamma$)\\
 \hline
 30  & 750 & 4,400   & 174.6 & 2  & 96.8 & 30  & 16.8 &  \cellcolor{red!20} 1.8 & \cellcolor{red!20} 10.4   \\
 125 & 750 & 4,400   & 174.6 & 9  & 21.5 & 125 & 4.0 &  \cellcolor{red!20} 8.1 & \cellcolor{red!20} 43.7  \\
 750 & 750 & 4,400   & 174.6 & 25 & 7.7  & 375 & 1.3 &  \cellcolor{red!20} 22.7 & \cellcolor{red!20} 134.3  \\

\hline
  \multicolumn{10}{c}{}\\ 
 \rowcolor{headcolor}
  \multicolumn{10}{c}{Experiment 4: Image Size}\\
  \hline
  \rowcolor{headcolor}
 image  &
 D (GB) & C (s) & D/C (MB/s) &
 $\gamma$ & $\delta/\gamma$ (MB/s) & $\Gamma$ & $\Delta/\Gamma$ (MB/s)&
(D/C)/($\delta/\gamma$) & (D/C)/($\Delta/\Gamma$)\\
 \hline
 BigBrain      & 750   & 2,200   & 349.1     & 9  & 21.5  & 125 & 4.0 &  \cellcolor{red!20} 16.2 & \cellcolor{red!20} 87.3   \\
 Half BigBrain & 375   & 2,200   & 174.6     & 9  & 21.5  & 125 & 4.0 &  \cellcolor{red!20} 8.1 & \cellcolor{red!20} 43.7   \\
 MRI           & 0.127 & 2,200   & 0.06      & 9  & 21.5  & 125 & 4.0 &  \cellcolor{green!20} 0.003 & \cellcolor{green!20} 0.015  
\end{tabular}
\caption{Experiment conditions. Red cells denote the conditions where 
the inequalities in Equations~\ref{eq:page-cache-inequality} 
and~\ref{eq:page-cache-sharedfs} do not hold, i.e., the page cache is 
asymptotically useless. Green cells show the conditions where the page cache covers all data writes.}
\label{table:experiments}
\end{table*}

@@

\begin{figure*}
\begin{subfigure}{\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{results/figures/iterations.pdf}
    \caption{Experiment 1: complete BigBrain, 125 chunks, 3.44-second tasks.}\label{fig:iterations}
\end{subfigure}
\begin{subfigure}{\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{results/figures/cputime.pdf}
    \caption{Experiment 2: complete BigBrain, 125 chunks, 10 iterations.}\label{fig:cputime}
\end{subfigure}
\begin{subfigure}{\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{results/figures/numchunks.pdf}
    \caption{Experiment 3: complete BigBrain, 10 iterations, C= 4,400s.}\label{fig:numchunks}
\end{subfigure}
\begin{subfigure}{\columnwidth}
    \centering
    \includegraphics[width=\linewidth]{results/figures/datasize.pdf}
    \caption{Experiment 4: 125 chunks, 10 iterations, 1.76-second tasks.}\label{fig:datasize}
\end{subfigure}
\caption{Experiment results: Makespans of Spark and Nipype writing to memory, tmpfs, local  disk and Lustre.}
\label{fig:results}
\end{figure*}
\subsection{Experiment 1: Number of Iterations}

@@

\subsection{Page Cache Model Evaluations}
\begin{figure*}
\begin{subfigure}{0.5\linewidth}
\centering
    \includegraphics[width=\textwidth]{results/figures/local-incrementation.pdf}
\caption{Local Disk}
\label{fig:modeleval-local}
    \end{subfigure}%
\begin{subfigure}{0.5\linewidth}
\centering
    \includegraphics[width=\textwidth]{results/figures/lustre-incrementation.pdf}
\caption{Lustre}
\label{fig:modeleval-lustre}
\end{subfigure}
    \caption{Model evaluation. Grey 
             regions denote areas that violate model predictions.}
\label{fig:modeleval}
\end{figure*}

In order to evaluate the page cache model, we compared the observed 
speedup ratio provided by in-memory computing to the (D/C) / 
($\delta$/$\gamma$) and (D/C) / ($\Delta$/$\Gamma$) ratios 
(Fig.~\ref{fig:modeleval}). Speed-up ratios were computed as the ratio 
between the makespan obtained with Spark on local disk or Lustre, and 
the makespan obtained with Spark for in-memory computing. Experiments 
for which there was no in-memory equivalent (i.e. BigBrain split into 
30 
chunks) 
were not considered.

@@

\section{Discussion} % 2 pages
\label{sec:discussion}

\subsection{Effect of In-Memory Computing}
% Shared fs vs local disk vs in-memory: what do we gain?
We measure the effect of in-memory computing by comparing the runs 
of Spark in-memory (yellow bars in Fig.~\ref{fig:results}) to the ones 
of Spark on local disk (non-hatched green bars). The speed-up provided 
by in-memory computing is also reported in 
Fig.~\ref{fig:modeleval-local}. The speed-up provided by in-memory 
computing increases with (D/C) / ($\delta$/$\gamma$), as expected from 
the model. In our experiments, it peaked at 1.6, for a ratio of 16.2. 
This correspond to the processing of the BigBrain with 125 chunks and 
1.76-second tasks in experiment 4 (total computing time C=2,200s), 
which is typically encountered in common image processing tasks such as 
denoising, intensity normalization, etc. The speed up of 1.6 is also 
reached with a ratio of 22.7 in experiment 3, obtained by processing 
the BigBrain with 750 chunks.

The results also allow us to speculate on the effect of in-memory 
computing on the pre-processing of functional MRI, another typical use 
case in neuroimaging. Assuming an average processing time of 20 minutes 
per subject, which is a ballpark value commonly observed with the 
popular SPM or FSL packages \todo{add refs}, an input data size of 100MB per subject, 
and an output data size of 2GB (20-fold increase compared to input 
size), the D/C ratio would be 1.8MB/s, which would reach the 
$\delta/\gamma$ threshold measured on this cluster for $\gamma=108$, 
that is, if 108 subjects were processed on the same node. This is very 
unlikely as the number of CPUs per node is 40. We therefore conclude 
that in-memory computing is likely to be useless for fMRI analysis.

%~ \subsection{Effect of Data Locality}
%~ % It should be important, in particular when there is contention and when
%~ % tasks are small. 
%~ We measure the effect of data locality by comparing the runs of Spark 
%~ on local disk (non-hatched green bars in Fig.~\ref{fig:results}) to the 
%~ ones of Spark on Lustre (non-hatched blue bars). The speed-up provided by 
%~ local execution peaks at 3.2, reached at 750 chunks in experiment 3. \todo{This needs to be finished or removed.}

\subsection{Combined Effect of In-Memory and Data Locality}
We measure the combined effect of data locality and in-memory computing 
by comparing the runs of Spark in-memory (yellow bars in 
Fig.~\ref{fig:results}) to the ones of Spark on Lustre (non-hatched 
blue bars). The speed-up provided by the combine use of data locality 
and in-memory computing is also reported in 
Fig.~\ref{fig:modeleval-lustre}. The provided speed-up increases with 
(D/C) / ($\Delta$/$\Gamma$), as expected from the model. In our 
experiments, it peaked around 5, for ratios of 120.4 and 64. Again, 
this configuration is likely to happen in typical image processing 
tasks performed on the BigBrain.

As for the fMRI speculation, the D/C 
ratio of 1.8MB/s would reach the $\Delta/\Gamma$ threshold for 
$\Gamma=280$, which is a realistic number of subjects to process on a 
complete cluster. Naturally, this estimate is highly dependent on the observed bandwidth
of the shared file system ($\Delta$).
